Synthesizing Symmetric Lenses

### 2018-06-30 papers/SymmetricLenses/main.tex
woo get in changes

\begin{figure}
  % https://tex.stackexchange.com/questions/168741/showing-two-listings-in-a-table-side-by-side
  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
0,20,40 cp -r /Library/ "$\$$(date)"

* df -h | sendmail admin@domain.com
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingA}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
<Tasks>
 <Task>
  <Command>c</Command>
  <Time>0</Time><Time>20</Time> <Time>40</Time>
 </Task>
 <Task>
  <Command>c</Command>
  <AllTimes/>
 </Task>
<Tasks>
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingB}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let minute_list = minute . ("," . minute)$^*$ in
let times = "*" | minute_list in
let line = times . " " . command in
let crontab = "" | line . ("\n\n" line)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingC}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let minute_xml = "<Time>" . minute . "</Time>" in
let minutes_xml = minute_xml . minute_xml$^*$ in
let xml_time = "<AllTimes/>" | minutes_xml in
let command_xml = 
  "<Command>" . command . "</Command>"
in
let xml_task = command_xml . xml_time in
let scheduled_job = "<Tasks>" . xml_task$^*$ . "</Tasks>"
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
      \hline
      crontab & scheduled.job \\
      \hline
      \listingA & \listingB \\
      \hline
      \listingC & \listingD \\
      \hline
    \end{tabular}
  }
  \caption{An example crontab and Scheduled Tasks file, and regular expressions
    describing the formats. For simplicity, these are modified versions of the
    real formats.}
  \label{fig:minimized-representations}
\end{figure}

We motivate our problem with an extended example. Consider the previously
described goal of keeping Linux cron jobs (crontab) and Windows Scheduled Tasks
(scheduled.job) synchronized. These are each formats specifying when to run
recurring tasks. For the sake of exposition, we use simplified versions crontab
and scheduled.job formats, with examples and corresponding regular expressions
shown in Figure~\ref{fig:minimized-representations}. Each file contains a list
of tasks set to run at specific times every hour. In both formats, users
can either provide specific minutes at which they want the program to run, or they
can specify that the program should run at every minute.

We wish for the jobs to be synchronized across our Windows and Linux devices.
However, the commands that need to be run are typically different for each
device. Paths are different in Windows and Linux, and, while there is typically a
Windows equvialent for a Linux command, that equivalent command is not always
named the same thing. So, while we want to be able to have the same number of
commands, with the commands running at the same times, we don't want to
align the actual commands that are run.

Lenses synchronize files by providing 4 functions, two \emph{create} functions,
and two \emph{put} functions. Create functions are used to initialize data in
one format when provided with data in the other. If a crontab is initialized, a
synchronized Scheduled Tasks file should also be created, and vice versa. Put
functions are used to propagate edits from one format to the other. If a crontab
is updated, we should be able to combine it with a non-updated Scheduled Tasks
file, and output a new Scheduled Tasks file, synchronized to the updated
crontab. To have a simple symmetric lens is to have these required
synchronization functions. Formally, a simple symmetric lens between $X$ and $Y$
comprises 4 functions:
\begin{enumerate}
\item $\CreateR \OfType X \rightarrow Y$
\item $\CreateL \OfType Y \rightarrow X$
\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
\item $\PutL \OfType Y \rightarrow X \rightarrow X$
\end{enumerate}
Furthermore, if a file is not edited, but is then resynchronized using a put
function, no data should be changed. To be a simple symmetric lens, one must
satisfy the following ``round-tripping'' laws, guaranteeing identity edits won't
change the file data.
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}

For synchronization, when the original crontab or Scheduled Tasks file is
created, a synchronized file of the other form is created from it with
$\CreateR$ or $\CreateL$ (depending on which file was initially created). On
future edits, changes are made with the puts, a \PutR would propagate the shared
information from the crontab, and knit it together with the information specific
to the Scheduled Tasks file.

\subsection{Programming Symmetric Lenses} We now build up a symmetric lens by
example. Our symmetric lens expressions are typed between regular expression
pairs, if \lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$}, then $\Lens$
encodes a simple symmetric lens between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$. At the end, we want our symmetric lenses to by typed
between \lstinline{crontab} and \lstinline{scheduled_job}.

The identity lens takes, as argument, a regular expression, and performs the
identity function on strings in that regular expression. The identity lens fully
aligns the data formats: when one data format is updated, the other format gets
the exact same value. In our example, when a job is told to run at a specific
time in crontab, it is also expected to run at that same time in the Scheduled
Jobs list. The identity lens propagates that information.
%
\begin{lstlisting}
id(minute) : minute $\Leftrightarrow$ minute
\end{lstlisting}
%

Where the identity lens aligns data, the disconnect lens does the opposite: it
leaves data unaligned. The disconnect lens takes 4 arguments: two regular
expressions and two strings. On creates, because there is no association between
data in the first format and data in the second format, designated default
strings are used for the other format. In our extended example, the Linux
commands should not be synchronized with the Windows commands, so a disconnect
lens would be used. We want our default behavior to be failures, under the
assumption this failure will raise a warning, in case somebody forgets to find
the analogous command of the other operating system.
%
\begin{lstlisting}
disc(command,command_xml,"false","<Command>SET ERRORLEVEL=1</Command>") 
  : command $\Leftrightarrow$ command_xml
\end{lstlisting}
%
When building a disconnect lens between two strings, one of which is merely the
empty string, the lenses \lstinline{ins} and \lstinline{del} can instead be
used, where \lstinline{ins($t$) = disc("",$t$,"",$t$)} and
\lstinline{del($s$) = disc($s$,"",$s$,"")}.

Finally, there are a number of lenses that compositionally build up new lenses
from old ones. For example, data can appear in one form or another: a program
can run during a specific set of minutes, or that program can run the every
minute. If \lstinline{minutes_map} and \lstinline{alltimes_map} are lenses that
synchronize those settings in each format, then those lenses can be composed
with an \OrLens lens.
%
\begin{lstlisting}
or(minutes_map,alltimes_map) : times $\Leftrightarrow$ xml_time
\end{lstlisting}

Data can also be concatenated, and so lens mappings can be concatenated.
However, synchronizing concatenated data can happen in a number of ways, as data
can appear in the same order, and data can be reordered. For example, in the
crontab, time data appears before command data, whereas in the Scheduled Jobs
file, command data comes before time data. If \lstinline{command_map} provides a
lens between commands in each format, and \lstinline{time_map} provides a lens
between time data in each format, those lenses can be composed with a \SwapLens
lens.
%
\begin{lstlisting}
swap(command_map,time_map) : line $\Leftrightarrow$ xml_task
\end{lstlisting}
%
Two lenses, $\Lens_1$ and $\Lens_2$, can be concatenated as
\lstinline{concat($\Lens_1$,$\Lens_2$)} or as the infix style
\lstinline{$\Lens_1$ . $\Lens_2$} for data that is not reordered.

Lastly, the \IterateLens lens synchronizes repeated data, aligning the number of
times the data is repeated. The \IterateLens combinator takes a lens as an
argument and synchronizes each element of the iteration with that lens. For
example, if \lstinline{full_task_map} synchronizes crontab lines prepended by
two newlines and xml for a single task, iterating that lens builds a lens that
synchronizes a list of tasks.
%
\begin{lstlisting}
iterate(full_task_map) : ("\n\n" . line)$^*$ $\Leftrightarrow$ xml_task$^*$
\end{lstlisting}
%

\begin{figure}
\begin{lstlisting}
let full_minute_map = ins("<Time>") . id(minute) . ins("</Time>") in
let comma_minute_map = del(",") . full_minute_map in
let minutes_map = full_minute_map . iterate(comma_minute_map) in
let alltimes_map = del("*") . ins("<AllTimes/>") in
let times_map = or(minutes_map,alltimes_map) in

let command_core = 
  disc(command,command_xml,"false","<Command>SET ERRORLEVEL=1</Command>")
in
let command_map = 
  ins("<Command>") . command_core . ins("</Command>")) . del(" ") 
in

let task_map = swap(command_map,time_map) in
let full_task_map = 
  ins("<Task>") . del("\n\n") . task_map . ins("</Task>") 
in
let repeated_task_map = full_task_map . iterate(full_task_map)
let nonempty_task_map = ins("<Tasks>") . repeated_task_map . ins("</Tasks>")

or(ins("<Tasks></Tasks>"),nonempty_task_map) : crontab $\Leftrightarrow$ scheduled_job
\end{lstlisting}
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes crontab files and Scheduled Task files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions. \saz{Say something
      about the possibility of synthesizing this lens?}}
  \label{fig:example_lens}
\end{figure}

Using these combinators, we can build up a lens that synchronizes crontabs and
Scheduled Task files, shown in Figure~\ref{fig:example_lens}. This lens uses one
typing rule not previously mentioned: type equivalence. While
\lstinline{scheduled_job} is a equivalent to a regular expression with an
outermost disjunction, it is not syntactically equivalent to one. Without type
equivalence, the provided lens wouldn't actually be well-typed, but type
equivalence states that if a lens is typed between a pair of regular
expressions, it is also typed between an equivalent pair.

@@

To alleviate these difficulties, we wish to synthesize symmetric lenses---given
a pair of regular expressions and a set of input-output examples, we want to
find a symmetric lens expression that is typed between the regular expression
pair, and satisfies the input/output examples. In this example, we wish to
synthesize a lens from \lstinline{crontab} and \lstinline{scheduled_job}, using
the strings in Figure~\ref{fig:minimized-representations} as an input-output
example for \CreateR.

However, our simple symmetric lens combinators permit many well-typed programs
between two regular expressions. For example, one possible lens between
\lstinline{crontab} and \lstinline{scheduled_job} is the one in
Figure~\ref{fig:example_lens}, but the lens
%
\begin{lstlisting}
disc(crontab,scheduled_job,rep(crontab),rep(scheduled_job))
\end{lstlisting}
%
is also a valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the synchronized files from
Figure~\ref{fig:minimized-representations} as an input-output example helps
little. The default strings given to \Disconnect merely need to be the strings
in the input-output examples, and the lens will still satisfy the specification.
While the disconnect lens is prohibited by two examples, or by a carefully
chosen input-output example on \PutR or \PutL, this problem continues presenting
itself for the sublenses, and is highlighted even more on larger formats. While
we could merely ask programmers to provide large, tight specifications for their
synthesis tasks, we would prefer the algorithm to choose the right lens
automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
synthesize the \emph{best} satisfying lens.

@@

Now that we have a notion of best, we need a way to
search for that best lens.
Our overall approach involves two communicating synthesizers: one (\RXSearch)
that proposes regular expression pairs, semantically but not syntactically
equivalent to the originals; one (\GreedySynth) that tries to find the best lens
within a class of lenses, characterized by such regular expression pairs. Given
two regular expressions, \GreedySynth can quickly search through the class of
lenses that can be typed between them, where all instances of type equivalence
involve no star equivalences. For example, consider the lens shown in
Figure~\ref{fig:example_lens}. This lens would \emph{not} be in the class
characterized by \lstinline{crontab} and \lstinline{scheduled_job}, as that
typing derivation requires unrolling \lstinline{xml_task$^*$} within
\lstinline{scheduled_job} into \lstinline{"" | (xml_task . xml_task$^*$)}.

@@

Our strategy involving communicating synthesizers permits both a way of
enumerating classes and efficiently searching through them, but it poses a
problem: when do we stop? While we may have found a great lens between two regular
expressions, a different regular expression pair may provide an even better
lens. Moreover, we may find a great lens characterized by
\lstinline{crontab} and \lstinline{scheduled_job} that is not the
desired lens shown in Figure~\ref{fig:example_lens}.  The
tension is that we want to find a good lens, but we also want to give quick feedback to
the user. After finding the lens in Figure~\ref{fig:example_lens}, we dont want
the algorithm to spend much longer searching for better lenses.

@@

\[
  \centering
  \inferrule*
  {
    \String \in \LanguageOf{\Regex}\\
    \StringAlt \in \LanguageOf{\RegexAlt}
  }
  {
    \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}
    \OfType \Regex \Leftrightarrow \RegexAlt
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String'$ & = & $\StringAlt$\\
    $\CreateL{} \App \StringAlt'$ & = & $\String$\\
    $\PutR{} \App \String' \App \StringAlt'$ & = & $\StringAlt'$\\
    $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
  \end{tabular}
\end{center}
Just as the identity lens ignores the second argument in puts, disconnect lenses
ignore the first in both puts and creates.  The data is unsynchronized in these
two formats, information from one format doesn't impact the other.

>>>>

\begin{figure}
  % https://tex.stackexchange.com/questions/168741/showing-two-listings-in-a-table-side-by-side
  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
Jane Doe: 35000
John Public: 37500
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingA}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
lastname,firstname,company
Doe,Jane,Healthcare Inc.
Public,John,Insurance Co.
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingB}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let salary = number | "UNSET"
let emp_salary = name . " " . name . ": " salary
let emp_salaries = "" | emp_salary . ("\n" emp_salary)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingC}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNSET" in
let emp_row = name . "," . name "," company in
let header = "firstname,lastname,company" in
let emp_insurances = header . ("\n" . emp_row)*
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
      \hline
      company.mgmt & company.hr \\
      \hline
      \listingA & \listingB \\
      \hline
      \listingC & \listingD \\
      \hline
    \end{tabular}
  }
  \caption{Example files used by management and HR, and regular expressions
    describing the formats of these files.}
  \label{fig:minimized-representations}
\end{figure}

\afm{i commented out the previous version, because this one was quickly written
  up to give you something to work off. Feel free to delete it if it isnt
  helpful. } We motivate our problem with an extended example. In our
theoretical company, human resources and upper management both keep information
about employees. HR needs to know the names of the employees as well as their
insurance providers. Management needs to know the names of employees and their
salary. Two example files, and regular expressions describing the formats, are
shown in Figure~\ref{fig:minimized-representations}. These two files should be
synchronized: if management adds a new employee, ``Richard Roe: 32500'', a new
row should be added to HR's file, ``Roe,Richard,UNSET''. Furthermore, if HR
changes ``John Public'' to ``Jon Public'', that name change should be reflected
in both files. However, not all the data should be synchronized. The HR file
should not get altered due to salary changes, and the management file should not
get altered due to insurance company changes. While names should be matched
across the two formats, insurance company and salary information should be left
unmatched. Such a synchronization can be achieved with simple symmetric lenses.

%Consider the previously
%described goal of keeping Linux cron jobs (crontab) and Windows Scheduled Tasks
%(scheduled.job) synchronized. These are each formats specifying when to run
%recurring tasks. For the sake of exposition, we use simplified versions crontab
%and scheduled.job formats, with examples and corresponding regular expressions
%shown in Figure~\ref{fig:minimized-representations}. Each file contains a list
%of tasks set to run at specific times every hour. In both formats, users
%can either provide specific minutes at which they want the program to run, or they
%can specify that the program should run at every minute.
%
%We wish for the jobs to be synchronized across our Windows and Linux devices.
%However, the commands that need to be run are typically different for each
%device. Paths are different in Windows and Linux, and, while there is typically a
%Windows equvialent for a Linux command, that equivalent command is not always
%named the same thing. So, while we want to be able to have the same number of
%commands, with the commands running at the same times, we don't want to
%align the actual commands that are run.

Lenses synchronize files by providing 4 functions, two \emph{create} functions,
and two \emph{put} functions. Create functions are used to initialize data in
one format when provided with data in the other. If a company.mgmt file is
initialized, a synchronized company.hr file should also be created, and vice
versa. Put functions are used to propagate edits from one format to the other.
If a company.hr file is updated, we should be able to combine it with a
non-updated company.mgmt file, and output a new company.mgmt file, synchronized
to the updated company.hr. To have a simple symmetric lens is to have these
required synchronization functions. Formally, a simple symmetric lens between
$X$ and $Y$ comprises 4 functions:
\begin{enumerate}
\item $\CreateR \OfType X \rightarrow Y$
\item $\CreateL \OfType Y \rightarrow X$
\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
\item $\PutL \OfType Y \rightarrow X \rightarrow X$
\end{enumerate}
Furthermore, if a file is not edited, but is then resynchronized using a put
function, no data should be changed. To be a simple symmetric lens, one must
satisfy the following ``round-tripping'' laws, guaranteeing identity edits won't
change the file data.
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{gather}

For synchronization, when the original company.hr or company.mgmt file is
created, a synchronized file of the other form is created from it with
$\CreateR$ or $\CreateL$ (depending on which file was initially created). On
future edits, changes are made with the puts, a \PutR would propagate the shared
information from the company.hr file, and knit it together with the information
specific to the company.mgmt file.

\subsection{Programming Symmetric Lenses} We now build up a symmetric lens by
example. Our symmetric lens expressions are typed between regular expression
pairs, if \lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$}, then $\Lens$
encodes a simple symmetric lens between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$. At the end, we want our symmetric lenses to by typed
between \lstinline{emp_salaries} and \lstinline{emp_insurance}.  The finished
lens is shown in Figure~\ref{fig:example_lens}.

The identity lens takes, as argument, a regular expression, and performs the
identity function on strings in that regular expression. The identity lens fully
matches the data formats: when one data format is updated, the other format gets
the exact same value. In our example, when a name is updated, it is also
expected to updated in the other format. The identity lens propagates that
information.
%
\begin{lstlisting}
id(name) : name $\Leftrightarrow$ name
\end{lstlisting}
%

Where the identity lens aligns data, the disconnect lens does the opposite: it
leaves data unaligned. The disconnect lens takes 4 arguments: two regular
expressions and two strings. On creates, because there is no association between
data in the first format and data in the second format, designated default
strings are used for the other format. In our extended example, salaries and
insurance companies should not be matched, so a disconnect lens would be used.
We want our default behavior to specify that the value has not been set.
%
\begin{lstlisting}
disc(salary,company,"UNSET","UNSET") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
When building a disconnect lens between two strings, one of which is merely the
empty string, the lenses \lstinline{ins} and \lstinline{del} can instead be
used, where \lstinline{ins($t$) = disc("",$t$,"",$t$)} and
\lstinline{del($s$) = disc($s$,"",$s$,"")}.

Finally, there are a number of lenses that compositionally build up new lenses
from old ones. For example, data can appear in one form or another: a program
can run during a specific set of minutes, or that program can run the every
minute. If \lstinline{minutes_map} and \lstinline{alltimes_map} are lenses that
synchronize those settings in each format, then those lenses can be composed
with an \OrLens lens.
%
\begin{lstlisting}
or(id(""),multi_emp_map) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_row) . ("\n" emp_row)*
\end{lstlisting}

Data can also be concatenated, and so lens mappings can be concatenated.  This
is achieved with the concat map.
%
\begin{lstlisting}
concat(ins(header),multi_emp_map) : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%
Two lenses, $\Lens_1$ and $\Lens_2$, can be concatenated as
\lstinline{concat($\Lens_1$,$\Lens_2$)} or as the infix style
\lstinline{$\Lens_1$ . $\Lens_2$}.

Lastly, the \IterateLens lens synchronizes repeated data, aligning the number of
times the data is repeated. The \IterateLens combinator takes a lens as an
argument and synchronizes each element of the iteration with that lens. For
example, if \lstinline{single_emp_map} synchronizes a single employee's data,
iterating that lens builds a lens that synchronizes a list of tasks.
%
\begin{lstlisting}
iterate(id("\n") . single_emp_map) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_row)$^*$
\end{lstlisting}
%

\begin{figure}
\begin{lstlisting}
let name_map = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let single_emp_map = name_map . disc(salary,company,"UNSET","UNSET") in
let multi_emp_map = ins("\n") . single_emp_map . iterate(id("\n") . single_emp_map) in
let emp_map = or(id(""),multi_emp_map) in
ins(header) . multi_emp_map : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and hr employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

Using these combinators, we can build up a lens that synchronizes HR and
management employee files. This lens uses one typing rule not previously
mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
a regular expression with an outermost disjunction, it is not syntactically
equivalent to one. Without type equivalence, the provided lens wouldn't actually
be well-typed, but type equivalence states that if a lens is typed between a
pair of regular expressions, it is also typed between an equivalent pair.

@@

To alleviate these difficulties, we wish to synthesize symmetric lenses---given
a pair of regular expressions and a set of input-output examples, we want to
find a symmetric lens expression that is typed between the regular expression
pair, and satisfies the input/output examples. In this example, we wish to
synthesize a lens from \lstinline{emp_salaries} and \lstinline{emp_insurance},
using the strings in Figure~\ref{fig:minimized-representations} as an
input-output example for \CreateR.

However, our simple symmetric lens combinators permit many well-typed programs
between two regular expressions. For example, one possible lens between
\lstinline{emp_salaries} and \lstinline{emp_insurance} is the one in
Figure~\ref{fig:example_lens}, but the lens
%
\begin{lstlisting}
disc(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is also a valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the synchronized files from
Figure~\ref{fig:minimized-representations} as an input-output example helps
little. The default strings given to \Disconnect merely need to be the strings
in the input-output examples, and the lens will still satisfy the specification.
While the disconnect lens is prohibited by two examples, or by a carefully
chosen input-output example on \PutR or \PutL, this problem continues presenting
itself for the sublenses, and is highlighted even more on larger formats. While
we could merely ask programmers to provide large, tight specifications for their
synthesis tasks, we would prefer the algorithm to choose the right lens
automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
synthesize the \emph{best} satisfying lens.

@@

Now that we have a notion of best, we need a way to search for that best lens.
Our overall approach involves two communicating synthesizers: one (\RXSearch)
that proposes regular expression pairs, semantically but not syntactically
equivalent to the originals; one (\GreedySynth) that tries to find the best lens
within a class of lenses, characterized by such regular expression pairs. Given
two regular expressions, \GreedySynth can quickly search through the class of
lenses that can be typed between them, where all instances of type equivalence
involve no star equivalences. For example, consider the lens shown in
Figure~\ref{fig:example_lens}. This lens would \emph{not} be in the class
characterized by \lstinline{emp_salaries} and \lstinline{emp_insurance}, as that
typing derivation requires unrolling \lstinline{("\n" . emp_salary)$^*$} within
\lstinline{emp_salaries} into \lstinline{"" | (("\n" . emp_salary) . ("\n" . emp_salary)$^*$)}.

@@

Our strategy involving communicating synthesizers permits both a way of
enumerating classes and efficiently searching through them, but it poses a
problem: when do we stop? While we may have found a great lens between two regular
expressions, a different regular expression pair may provide an even better
lens. Moreover, we may find a great lens characterized by
\lstinline{emp_salaries} and \lstinline{emp_insurances} that is not the
desired lens shown in Figure~\ref{fig:example_lens}.  The
tension is that we want to find a good lens, but we also want to give quick feedback to
the user. After finding the lens in Figure~\ref{fig:example_lens}, we dont want
the algorithm to spend much longer searching for better lenses.

@@

\[
  \centering
  \inferrule*
  {
    \String \in \LanguageOf{\Regex}\\
    \StringAlt \in \LanguageOf{\RegexAlt}
  }
  {
    \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}
    \OfType \Regex \Leftrightarrow \RegexAlt
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String'$ & = & $\StringAlt$\\
    $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\
    $\PutR{} \App \String'' \App \StringAlt'$ & = & $\StringAlt'$\\
    $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
  \end{tabular}
\end{center}
Just as the identity lens ignores the second argument in puts, disconnect lenses
ignore the first in both puts and creates.  The data is unsynchronized in these
two formats, information from one format doesn't impact the other.
